---
title: "Final Project"
author:
- Ali Harb
- Dilip Ganesan
- Raghunathan Ramnath
date: "May 10, 2019"
output:
  pdf_document:
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  prettydoc::html_pretty:
    df_print: paged
    highlight: github
    theme: leonids
    toc: yes
geometry: margin=1in
link-citations: yes
fontsize: 11pt
subtitle: CUNY SPS Masters in Data Science - DATA 698
fontfamily: mathpazo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

withCallingHandlers(suppressWarnings(warning("hi")), warning = function(w) {
    print(w)
})

# Loading Packages
requiredpackages <- c('knitr','psych','nlme','lattice', 'kableExtra', 'psych', 'ggplot2', 'reshape2', 'corrplot', 'tidyr', 'dplyr', 'plyr', 'MASS', 'caret', 'pscl', 'lmtest', 'pROC', 'ROCR','tibble','leaps','MASS','magrittr','ggplot2','glmnet','faraway','choroplethr','gridExtra','grid','aod')

for (i in requiredpackages){
  if(!require(i,character.only=T)) install.packages(i)
  library(i,character.only=T)
}

update_geom_defaults("point", list(size=1.5))
theme_set(theme_grey(base_size=10))

```


\newpage
#Load cleaned Data

```{r P1, include=FALSE}
#Load data set
dset <- read.csv("https://raw.githubusercontent.com/dilipganesan/DATA698/master/Cleaned_Data/Cleaned_DATA698_Dataset.csv")
                 
head(dset)                 
```


```{r, echo=FALSE}
# Obtain dimensions of dataset
get_dims <- function(df){
  
  dimensions <- dim(df)
  dimensions <- data.frame('Records' = dimensions[1],
                           'Variables' = dimensions[2])
  return(dimensions)
}

```



```{r P2, echo=FALSE}

dimensions <- get_dims(dset)
```

The current data set is composed of `r dimensions$Records[1]` records and `r dimensions$Variables[1]` variables.


\newpage
# Summary Statistics


```{r P3, echo=FALSE}
summary(dset)
```

\newpage
# Data Analysis

## Histogram

```{r p6, echo=FALSE}

ggplot(melt(dset), aes(x=value)) + facet_wrap(~variable, scale="free") + geom_histogram(bins=50)

```

\newpage

## Missing Values: 
```{r p7, echo=FALSE}
#Replace ? with NA
#dset[dset=="?"]<-NA

## Missing Values
options(scipen = 999)
missing_plot <- VIM::aggr(dset,  
                      numbers = T, 
                      sortVars = T,
                      col = c("lightgreen", "darkred", "orange"),
                      labels=str_sub(names(dset), 1, 8), 
                      ylab=c("Missing Value Counts"
                             , "Pattern"))


missing_plot$missings %>% 
  mutate(
    pct_missing = Count / nrow(dset)
    ) %>% 
  arrange(-pct_missing) %>% 
  filter(pct_missing > 0) %>% 
  kable(digits = 3, row.names = T, caption = "Variables Missing Values")  
options(scipen=0, digits=7)

```

# Drop Missing Values: 

Weight is missing in over 98% records. Owing to the poor interpretability of missing values and little predictive generalizability to other patients, best thing is to just drop it.

Payer code and Medical Specialty of treating physician also have 40-50% missing values. We decided to drop these.

```{r p8, echo=FALSE}

dset$weight<-NULL
dset$payer_code<-NULL
dset$medical_specialty<-NULL
dset$citoglipton<-NULL
dset$examide<-NULL

#deletes columns 'weight','payer_code','medical_specialty'

#We also noticed that for two variables (drugs named citoglipton and examide), all records have the same value. So essentially these cannot provide any interpretive or discriminatory information for predicting readmission, and we dropped these columns as well. Technically, this isn't a missing value problem but rather a missing information problem.

```

# Addition of variables: 

  Service utilization: The data contains variables for number of inpatient (admissions), emergency room visits and outpatient visits for a given patient in the previous one year. These are (crude) measures of how much hospital/clinic services a person has used in the past year.
 
```{r p9, echo=FALSE}

dset['service_utilization'] <- dset['number_outpatient'] + 
   dset['number_emergency'] + dset['number_inpatient']

```

 Number of medication changes: The dataset contains 23 features for 23 drugs (or combos) which indicate for each of these, whether a change in that medication was made or not during the current hospital stay of patient. Medication change for diabetics upon admission has been shown by previous research to be associated with lower readmission rates. We decided to count how many changes were made in total for each patient, and declared that a new feature. The reasoning here was to both simplify the model and possibly discover a relationship with number of changes regardless of which drug was changed.




```{r p11, echo=FALSE}

dset$numchange<- dset['metformin'] + dset['repaglinide'] + dset['nateglinide'] + dset['chlorpropamide'] + dset['glimepiride']   + dset['glipizide']+ dset['glyburide']+ dset['pioglitazone']   + dset['rosiglitazone'] + dset['acarbose'] + dset['miglitol'] + dset['insulin'] + dset['glyburide.metformin']  + dset['tolazamide'] + dset['metformin.pioglitazone'] + dset['metformin.rosiglitazone']+ dset['glimepiride.pioglitazone'] + dset['glipizide.metformin'] + dset['troglitazone']+ dset['tolbutamide'] + dset['acetohexamide']



```

# Data Required for Modeling: 
```{r p12, echo=FALSE}
nset <-data.frame(
dset$race,dset$gender,	 dset$age,	dset$admission_type_id,	dset$discharge_disposition_id,	dset$admission_source_id,	dset$time_in_hospital,	dset$num_lab_procedures,	dset$num_procedures,	dset$num_medications,	dset$number_outpatient,	dset$number_emergency,	dset$number_inpatient,	dset$number_diagnoses,	dset$max_glu_serum,	dset$A1Cresult,dset$change,	dset$diabetesMed,	dset$readmitted,dset$DRG,dset$Payment,dset$primarydiagclass,	dset$secondarydiagclass_1,	dset$tertiarydiagclass_1,	dset$DRGClassification,
dset$service_utilization, dset$numchange)

colnames(nset) <-c("gender","age","admission_type_id","discharge_disposition_id",	"admission_source_id",	"time_in_hospital",	"num_lab_procedures",	"num_procedures",	"num_medications",	"number_outpatient",	"number_emergency",	"number_inpatient",	"number_diagnoses",	"max_glu_serum","A1Cresult","change",	"diabetesMed",	"readmitted","DRG","Payment","primarydiagclass",	"secondarydiagclass_1",	"tertiarydiagclass_1",	"DRGClassification",
"service_utilization", "numchange")


dimensions <- get_dims(nset)
```

The new data set is composed of `r dimensions$Records[1]` records and `r dimensions$Variables[1]` variables.
  


# Split Data: 
```{r p13, echo=FALSE}
    
    train<-sample_frac(nset, 0.7)
    sid<-as.numeric(rownames(train)) # because rownames() returns character
    test<-nset[-sid,]
    
```


# Model1: Binary Logistic Regression Model with no DRG

```{r p14, echo=FALSE}

#withCallingHandlers(suppressWarnings(warning("hi")), warning = function(w) {
 #   print(w)
#})

model1_no_drg <- train %>%dplyr::select(-train$DRG)
model1_lg <- glm(train$readmitted ~ ., family="binomial", model1_no_drg)

summary(model1_lg)

# Checking residual plot for model number 1
par(mfrow=c(2,2))
plot(model1_lg)

```

coefficient is negative for gender, admission_type_id, admission_source_id,
num_lab_procedures,num_medications, diabetesMed,       primarydiagclass,secondarydiagclass_1,DRClassification and service_utilization.

standard error is high for gender, discharge_disposition_id,    num_medications,number_emergency,Payment,number_diagnoses.


# CIs using standard errors

```{r p14.2, echo=FALSE}

#confidence interval using profiled log-likelihood
#confint(model1_lg)

## CIs using standard errors
confint.default(model1_lg)

```



We can test for an overall effect of rank using the wald.test function of the aod library. The order in which the coefficients are given in the table of coefficients is the same as the order of the terms in the model. This is important because the wald.test function refers to the coefficients by their order in the model. We use the wald.test function. b supplies the coefficients, while Sigma supplies the variance covariance matrix of the error terms, finally Terms tells R which terms in the model are to be tested, in this case, terms 1 to 13 and 24 to 26.

# Odds Ratio

```{r p14.1, echo=FALSE}
library(lmtest)
#waldtest(model1_lg, "factor(base_repo_id)", test = "Chisq")  
#l <- cbind(0, 0, 0, 1, -1, 0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0,0)
#wald.test(b = coef(model1_lg), Sigma = vcov(model1_lg), L = l)

## odds ratios only
exp(coef(model1_lg))

## odds ratios and 95% CI
#exp(cbind(OR = coef(model1_lg), confint(model1_lg)))


```


# Model1: Binary Logistic Regression Model with DRG

```{r p15, echo=FALSE}

model1_drg <- train 
model1_lg <- glm(train$readmitted ~ ., family="binomial", model1_drg)

summary(model1_lg)

# Checking residual plot for model number 1 with DRG
par(mfrow=c(2,2))
plot(model1_lg)

```

AIC is slightly more with DRG than the model with no DRG.