---
title: "Final Project"
author:
- Ali Harb
- Dilip Ganesan
- Raghunathan Ramnath
date: "May 10, 2019"
output:
  pdf_document:
    df_print: kable
    fig_caption: yes
    fig_height: 6
    fig_width: 7
    highlight: tango
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 4
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '4'
  prettydoc::html_pretty:
    df_print: paged
    highlight: github
    theme: leonids
    toc: yes
geometry: margin=1in
link-citations: yes
fontsize: 11pt
subtitle: CUNY SPS Masters in Data Science - DATA 698
fontfamily: mathpazo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE, warning=FALSE, error=FALSE, cache=FALSE, results='hide', message=FALSE}

library(kableExtra) # latex tables
library(stringr)    # strings
library(psych)      # describe a data frame
#library(zipcode)    # zipcode library
library(dplyr)      # data wrangling
library(tidyr)      # data wrangling
library(reshape2)   # data wrangling
library(nlme)       # generalized linear model
library(lattice)    # plotting
#library(fastDummies)# dummy variables
library(ggplot2)    # graphical plots
#library(ggpubr)     # plots
library(grid)       # map plots
library(gridExtra)
#library(choroplethrZip)
library(choroplethr)

```

#Load Data

```{r P1}
#Load data set
dset <- read.csv("C:/cuny/2019/698/dataset_diabetes/dataset_diabetes/diabetic_data.csv")
                 
head(dset)                 
```


```{r, echo=FALSE}
# Obtain dimensions of dataset
get_dims <- function(df){
  
  dimensions <- dim(df)
  dimensions <- data.frame('Records' = dimensions[1],
                           'Variables' = dimensions[2])
  return(dimensions)
}

```



```{r P2, echo=FALSE}

dimensions <- get_dims(dset)
```

\newpage
# Summary Statistics

The current data set is composed of `r dimensions$Records[1]` records and `r dimensions$Variables[1]` variables.

```{r P3, echo=FALSE}

#dset$admission_type_id <- factor(dset$admission_type_id)
#dset$discharge_disposition_id <- factor(dset$discharge_disposition_id)
#dset$admission_source_id <- factor(dset$admission_source_id)
#dset$time_in_hospital<-factor(dset$time_in_hospital)
#dset$num_lab_procedures<-factor(dset$num_lab_procedures)
#dset$num_procedures<-factor(dset$num_procedures)
#dset$num_medications<-factor(dset$num_medications)
#dset$number_outpatient<-factor(dset$number_outpatient)
#dset$number_emergency<-factor(dset$number_emergency)
#dset$number_inpatient<-factor(dset$number_inpatient)
#dset$number_diagnoses<-factor(dset$number_diagnoses)


summary(dset)
```


#Data Analysis

encounter_id

patient_nbr

Race: 75% of the population are Caucasian. Remaining 25% includes other 4 races that include others and unknown.

Gender: Almost 54% of the population is female and 46% of the population is male.

Age: [70-80] has the highest patient population.

admission_type_id: Almost 54% of the patient population were admitted in 
"Emergency"

discharge_disposition_id:60% of the patient population where discharged to home

admission_source_id: "Emergency Room" was the source for 57% of patient population

payer_code:40% of the patient population does not have payer_code. 

medical_specialty:50% of the population does not have the specialty. 

num_lab_procedures:varies..

num_procedures:47% patient population did not have any procedures.  21% had 1 procedure and 13% had 2 procedures and the remaining had between 3 and 6.

num_medications: varies..

number_outpatient:85% its 0

number_emergency:90% of time its not emergency

number_inpatient:67% its 0.

number_diagnoses:50% of the population have 9 diagnoses

max_glue_serum:96% of the patient population does not have.

A1Cresult:85% of the patient population did not have A1C.

metformin:81% of the patient population did not have metformin.

insulin:varies

change:47% its ch.

diabetesMed:78% its yes.

readmitted:46% admitted.

diag_1

diag_2

diag_3


## Following can be ignored.

weight: this can be ignored as for 98% of the patient population its unknown.

repaglinide:Mostly its no.

naateglinide:Mostly its no.

chlorpropamide:Mostly its no.

glimepiride:Mostly its no.

acetohexamide:No except one.

glipizide:No except one.

glyburide:91% its no.

miglitol:91% its no.

troglitazone:Mostly its no.

tolazamide:Mostly its no.

examide:Mostly its no.

citoglipton:Mostly its no.

glyburide.metformin:Mostly its no.

glipizide.metformin:Mostly its no.

glimepiride.pioglitazone:Mostly its no.

```{r P4, echo=FALSE}
str(dset)
```

# Histogram

```{r p6}

ggplot(melt(dset), aes(x=value)) + facet_wrap(~variable, scale="free") + geom_histogram(bins=50)

```

# Missing Values: 
```{r p7}
#Replace ? with NA
dset[dset=="?"]<-NA

## Missing Values
options(scipen = 999)
missing_plot <- VIM::aggr(dset,  
                      numbers = T, 
                      sortVars = T,
                      col = c("lightgreen", "darkred", "orange"),
                      labels=str_sub(names(dset), 1, 8), 
                      ylab=c("Missing Value Counts"
                             , "Pattern"))


missing_plot$missings %>% 
  mutate(
    pct_missing = Count / nrow(dset)
    ) %>% 
  arrange(-pct_missing) %>% 
  filter(pct_missing > 0) %>% 
  kable(digits = 3, row.names = T, caption = "Variables Missing Values")  
options(scipen=0, digits=7)

```

# Drop Missing Values: 

Weight is missing in over 98% records. Owing to the poor interpretability of missing values and little predictive generalizability to other patients, best thing is to just drop it.

Payer code and Medical Specialty of treating physician also have 40-50% missing values. We decided to drop these.

```{r p8}

dset$weight<-NULL
dset$payer_code<-NULL
dset$medical_specialty<-NULL
dset$citoglipton<-NULL
dset$examide<-NULL

#deletes columns 'weight','payer_code','medical_specialty'

#We also noticed that for two variables (drugs named citoglipton and examide), all records have the same value. So essentially these cannot provide any interpretive or discriminatory information for predicting readmission, and we dropped these columns as well. Technically, this isn't a missing value problem but rather a missing information problem.

```

# Addition of variables: 

 Service utilization: The data contains variables for number of inpatient (admissions), emergency room visits and outpatient visits for a given patient in the previous one year. These are (crude) measures of how much hospital/clinic services a person has used in the past year.
 
```{r p9}

dset['service_utilization'] <- dset['number_outpatient'] + 
   dset['number_emergency'] + dset['number_inpatient']

```

Number of medication changes: The dataset contains 23 features for 23 drugs (or combos) which indicate for each of these, whether a change in that medication was made or not during the current hospital stay of patient. Medication change for diabetics upon admission has been shown by previous research to be associated with lower readmission rates. We decided to count how many changes were made in total for each patient, and declared that a new feature. The reasoning here was to both simplify the model and possibly discover a relationship with number of changes regardless of which drug was changed.


 
```{r p10}

keys = list ('metformin', 'repaglinide', 'nateglinide', 'chlorpropamide',
'glimepiride', 'glipizide', 'glyburide', 'pioglitazone', 'rosiglitazone',
 'acarbose', 'miglitol', 'insulin', 'glyburide.metformin', 'tolazamide', 
'metformin.pioglitazone','metformin.rosiglitazone', 'glimepiride.pioglitazone', 
'glipizide.metformin', 'troglitazone', 'tolbutamide', 'acetohexamide')


bv <- function(x) {  
    if (x =='No'|x=='Steady'){s=0}
      else {s=1}
    return(s)
}

#dset$metformin <- dset$metformin.apply(lambda x: 0 if (x =='No'|x=='Steady') else 1)
#dset['metformin']<- sapply(dset['metformin'], bv)

for (i in keys)
  { dset[i]<- sapply(dset[i], bv)
}
```


```{r p11}

dset$numchange<- dset['metformin'] + dset['repaglinide'] + dset['nateglinide'] + dset['chlorpropamide'] + dset['glimepiride']   + dset['glipizide']+ dset['glyburide']+ dset['pioglitazone']   + dset['rosiglitazone'] + dset['acarbose'] + dset['miglitol'] + dset['insulin'] + dset['glyburide.metformin']  + dset['tolazamide'] + dset['metformin.pioglitazone'] + dset['metformin.rosiglitazone']+ dset['glimepiride.pioglitazone'] + dset['glipizide.metformin'] + dset['troglitazone']+ dset['tolbutamide'] + dset['acetohexamide']

#for (i in keys)
#    { dset['numchange'] = dset['numchange'] + dset[i] }

```


# Split Data: 
```{r p12}
    
    train<-sample_frac(dset, 0.7)
    sid<-as.numeric(rownames(train)) # because rownames() returns character
    test<-dset[-sid,]
    
```


