---
title: "Data Transformation"
author: "Ali Harb"
date: "March 2, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Import and Explore Data

### Import the diagnostic data to R IDE

```{r}
MyData <- read.csv(file="diag.csv", header=TRUE, sep=",")
MyData_str<- data.frame(lapply(MyData, as.character), stringsAsFactors=FALSE)
str(MyData_str)
head(MyData)
```

### Gather all the diagnostic codes into one vector 

```{r}
diags<-c(as.character(MyData$diag_1),as.character(MyData$diag_2),as.character(MyData$diag_3))
head(diags)
print(paste("The length of the vector: ",length(diags)))
```

### Get the number of the unique diagnostic codes

```{r}
UniqueDiags<-unique(diags)
print(paste("The number of unique codes: ",length(UniqueDiags)))
```


### Combine diagnostic codes of diag 1,2 and 3 into a single value as diagnostic sets to find how many combinations of codes the dataset contain

```{r}
diagSets=0
for( i in 0:nrow(MyData)){
  diagSets[i]=paste0(MyData$diag_1[i],'|',MyData$diag_2[i],'|',MyData$diag_3[i])
}
diagSets[0:10]

print(paste("The number of combinations: ",length(diagSets)))
```

### Get the number of unique combination sets of diagnostic codes

```{r}
unique_diagSets<-unique(diagSets)
print(paste("The number of unique combinations: ",length(unique_diagSets)))
```

### Find the letters which the diagnostic codes utilizes 

```{r}
for(i in LETTERS){
 x<-diagE_unique<-UniqueDiags[startsWith(UniqueDiags,i)]
 if(length(x) !=0)
 {
   print(i)
 }
}
```

## Tranform Data 

### Get all codes that start with V letter

```{r}
diagV_unique<-UniqueDiags[startsWith(UniqueDiags, "V")]
diagV_unique
cat("\n")
print(paste("The number of unique values: ",length(diagV_unique)))
```
Matched:
V08

Matched only if you add .01:
V53,V71,V54,V25,V15,V61

Add 00 to the end of the code:
V45,V67,V10,V12,V64,V13

Add 0 to the rest of the code

#### Modify the V-codes to match the Find-A-Code format

```{r}
diagNot<-c("V53","V71","V54","V25","V15","V61")
diag00<-c("V45","V67","V10","V12","V64","V13")
diagNochange<-c("V08")

diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-diagV_unique[!(diagV_unique %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-diagNot

# test the codes

unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "V")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "V")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "V")])
```


### Get all codes that start with E

```{r}
diagE_unique<-UniqueDiags[startsWith(UniqueDiags, "E")]
diagE_unique
cat("\n")
print(paste("The number of unique values: ",length(diagE_unique)))
```

Matched:
E956,E915,E918,E887,E916,E882,E894,E966,E892,E912

Add 0 at the end for the rest of the codes 


#### Modify the E-codes to match the Find-A-Code format

```{r}
diagNochange<-c("E956","E915","E918","E887","E916","E882","E894","E966","E892","E912")

diag0<-diagE_unique[!(diagE_unique %in% diagNochange)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

# test the codes

unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "E")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "E")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "E")])
```


### Get the single digit number 

```{r}
lessThan2<-diags[nchar(diags)==1]
unique(lessThan2)
cat("\n")
print(paste("The number of unique values: ",length(lessThan2)))

na<-lessThan2[startsWith(lessThan2, "?")]
cat("\n")
print(paste("The number of elements contain '?' : ",length(na)))

```

Add double zero
8 = 00800

all the rest add 00 before and 0 after


#### Replace '?' with "" 

```{r}
MyData_str[MyData_str=="?"]<-""
head(MyData_str)
```


#### Transform the single digit numbers to 3 or more digits codes to match the Find-A-Code format

```{r}
diag0<-c("3","7","5","9")

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0("00",i,"0")
}


MyData_str[MyData_str=="8"]<-paste0("00","8","00")

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "00")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "00")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "00")])
```


### Get the doule digit numbers 

```{r}
lessThan3<-diags[nchar(diags)==2]
ulessThan3<-unique(lessThan3)
ulessThan3
cat("\n")
print(paste("The number of unique values: ",length(ulessThan3)))
```
No Match:
58

Add 0 before and 00 after:
41,11,10,17,14

Add 0 before:
42,48,35,75,61,96

Add 0 before and 0 after for the rest of the code

#### Transform the double digit munbers to 3 or more digit codes to match the Find-A-Code format

```{r}
diagNot<-c("58")
diag0Before00<-as.character(c(41,11,10,17,14))
diag0Before<-as.character(c(42,48,35,75,61,96))


diagNot0<-c(diagNot,diag0Before00,diag0Before)

diag0<-ulessThan3[!(ulessThan3 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0("0",i,"0")
}

for(i in diag0Before00){
  MyData_str[MyData_str==i]<-paste0("0",i,"00")
}

for(i in diag0Before){
  MyData_str[MyData_str==i]<-paste0("0",i)
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)
# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "01")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "03")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "04")])
```


### Create a vector of 3 digit numbers codes excluding V-codes

```{r}
lessThan4<-diags[nchar(diags)==3]
lessThan4<-lessThan4[!startsWith(lessThan4, "V")]
unique(lessThan4)
cat("\n")
print(paste("The number of unique values: ",length(unique(lessThan4))))
```


### Get all 3 digit numbers that start with 1 from the new vector

```{r}
lessThan41<-lessThan4[startsWith(lessThan4, "1")]
ulessThan41<-unique(lessThan41)
ulessThan41
cat("\n")
print(paste("The number of unique values: ",length(ulessThan41)))

```
No Match: 
187

Matched:
185,135,193,179,138

Add 00 at the end of the code:
173,115,131

Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 1 to match the Find-A-Code format

```{r}
diagNot<-c("187")
diag00<-c("173","115","131")
diagNochange<-c("185","135","193","179","138")

diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan41[!(ulessThan41 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "1")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "1")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "1")])
```


### Get the 3 digit numbers that start with 2 from the vector 

```{r}
lessThan42<-lessThan4[startsWith(lessThan4, "2")]
ulessThan42<-unique(lessThan42)
ulessThan42
cat("\n")
print(paste("The number of unique values: ",length(ulessThan42)))

```

No Match:
284,275,258,260,243


Matched:
220,217,226,261,262


Add 00 at the end of the code:
277,295,296,202,200,288,250,204,278,203,201,274,228,242,205,250,207,208,299,279

Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 2 to match the Find-A-Code format

```{r}
diagNot<-c("284","275","258","260","243")
diag00<-c("277","295","296","202","200","288","250","204","278","203","201","274","228","242","205","250","207","208","299")
diagNochange<-c("220","217","226","261","262")

diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan42[!(ulessThan42 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "2")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "2")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "2")])
```

### Get the 3 digit numbers that Start with 3

```{r}
lessThan43<-lessThan4[startsWith(lessThan4, "3")]
ulessThan43<-unique(lessThan43)
ulessThan43
cat("\n")
print(paste("The number of unique values: ",length(ulessThan43)))

```

No Match:
350,312,362,323


Matched:
340,311,395,319,317,316


Add 00 at end of the code:
386,388,337,346,303,380,358,345,305,368,304,366,374,376,383,385,300,361,370,365,344,389,373,378,342,369,379,382,384,377,375,347,360,314,327,363,372,381,364,315

Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 3 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(350,312,362,323))
diag00<-as.character(c(386,388,337,346,303,380,358,345,305,368,304,366,374,376,383,385,300,361,370,365,344,389,373,378,342,369,379,382,384,377,375,347,360,314,327,363,372,381,364,315))
diagNochange<-as.character(c(340,311,395,319,317,316))

diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan43[!(ulessThan43 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "3")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "3")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "3")])
```

### Get the 3 digit numbers that Start with 4

```{r}
lessThan44<-lessThan4[startsWith(lessThan4, "4")]
ulessThan44<-unique(lessThan44)
ulessThan44
cat("\n")
print(paste("The number of unique values: ",length(ulessThan44)))
```
No Match:
444,445,405,484


Matched:
462,431,481,496,485,475,463,430,490,470,452,412,460


Add 00 at the end of the code:
414,434,410,402,486,433,493,441,404,436,464,474


Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 4 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(444,445,405,484))
diagNochange<-as.character(c(462,431,481,496,485,475,463,430,490,470,452,412,460))
diag00<-as.character(c(414,434,410,402,486,433,493,441,404,436,464,474))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan44[!(ulessThan44 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "4")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "4")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "4")])
```

### Get the 3 digit numbers that Start with 5

```{r}
lessThan45<-lessThan4[startsWith(lessThan4, "5")]
ulessThan45<-unique(lessThan45)
ulessThan45
cat("\n")
print(paste("The number of unique values: ",length(ulessThan45)))
```

No match:
584,558,585,517


Matched:
515,514,566,586,591,570,501,541,542,500,538


Add 00 at the end of the code:
564,590,574,562,553,531,532,535,550,598,552,534,533,519,528,551,521,524,523

Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 5 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(584,558,585,517))
diagNochange<-as.character(c(515,514,566,586,591,570,501,541,542,500,538))
diag00<-as.character(c(564,590,574,562,553,531,532,535,550,598,552,534,533,519,528,551,521,524,523))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan45[!(ulessThan45 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

# test the codes
nonMatchingCodes<-c(nonMatchingCodes,diagNot)
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "5")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "5")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "5")])
```


### Get the 3 digit numbers that Start with 6

```{r}
lessThan46<-lessThan4[startsWith(lessThan4, "6")]
ulessThan46<-unique(lessThan46)
ulessThan46
cat("\n")
print(paste("The number of unique values: ",length(ulessThan46)))
```

No Match:
645,690,624

Matched:
632,683,605


Add 00 ata the end of the code:
648,618,642,656,661,681,654,655,658,644,669,663,664,600,659,647,652,646,641,643,660,634,686,637,633,640,674,653,665,657,649,670


Add 0 at the end to rest of codes

#### Transform the 3 digit numbers that start with 6 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(645,690,624))
diagNochange<-as.character(c(632,683,605))
diag00<-as.character(c(648,618,642,656,661,681,654,655,658,644,669,663,664,600,659,647,652,646,641,643,660,634,686,637,633,640,674,653,665,657,649,670))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan46[!(ulessThan46 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "6")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "6")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "6")])
```


### Get the 3 digit numbers that Start with 7

```{r}
lessThan47<-lessThan4[startsWith(lessThan4, "7")]
ulessThan47<-unique(lessThan47)
ulessThan47
cat("\n")
print(paste("The number of unique values: ",length(ulessThan47)))
```

No Match:  (family code)
780,787,799,790 

Matched:
797,725,734


Add 00 ata the end of the code:
786,707,715,730,724,789,733,711,727,716,736,746,719,718,709,741,755,744

Add 0 at the end to rest of codes


#### Transform the 3 digit numbers that start with 7 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(780,787,799,790 ))
diagNochange<-as.character(c(797,725,734))
diag00<-as.character(c(786,707,715,730,724,789,733,711,727,716,736,746,719,718,709,741,755,744))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan47[!(ulessThan47 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "7")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "7")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "7")])
```


### Get the 3 digit numbers that Start with 8

```{r}
lessThan48<-lessThan4[startsWith(lessThan4, "8")]
ulessThan48<-unique(lessThan48)
ulessThan48
cat("\n")
print(paste("The number of unique values: ",length(ulessThan48)))
```

No Match:

Matched:


Add 00 zeros at the end of the code: 
845,812,807,820,823,852,805,851,803,821,868,813,961,801,853,810,864,800,831,839,815,814,865,816,835,881,866,890,834,854,838,880,842,806,804,832,833,872,811

Add 0 at the end to rest of codes


#### Transform the 3 digit numbers that start with 8 to match the Find-A-Code format

```{r}
#diagNot<-as.character(c())
#diagNochange<-as.character(c())
diag00<-as.character(c(845,812,807,820,823,852,805,851,803,821,868,813,961,801,853,810,864,800,831,839,815,814,865,816,835,881,866,890,834,854,838,880,842,806,804,832,833,872,811))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan48[!(ulessThan48 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "8")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "8")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "8")])
```


### Get the 3 digit numbers that Start with 9

```{r}
lessThan49<-lessThan4[startsWith(lessThan4, "9")]
ulessThan49<-unique(lessThan49)
ulessThan49
cat("\n")
print(paste("The number of unique values: ",length(ulessThan49)))
```

No match:
959 

Matched:
936,920,986


Add 00 at the end of the code:
996,998,997,969,965,924,942,989,945,928,952,944,923,941,903,948,927,943

Add 0 at the end to the rest of the codes


#### Transform the 3 digit numbers that start with 9 to match the Find-A-Code format

```{r}
diagNot<-as.character(c(959))
diagNochange<-as.character(c(936,920,986))
diag00<-as.character(c(996,998,997,969,965,924,942,989,945,928,952,944,923,941,903,948,927,943))


diagNot0<-c(diagNot,diag00,diagNochange)

diag0<-ulessThan49[!(ulessThan49 %in% diagNot0)]

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

for(i in diag00){
  MyData_str[MyData_str==i]<-paste0(i,"00")
}

nonMatchingCodes<-c(nonMatchingCodes,diagNot)

# testing codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, "9")])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, "9")])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, "9")])
```

### Get the 4 digit numbers excluding the E-code

```{r}
lessThan5<-diags[nchar(diags)==4]
lessThan5<-lessThan5[!startsWith(lessThan5, "E")]
unique(lessThan5)
cat("\n")
print(paste("The number of unique values: ",length(unique(lessThan5))))
```

Zero 4 digit numbers 

### Get the 5 digit numbers

```{r}
lessThan6<-diags[nchar(diags)==5]
unique(lessThan6)
cat("\n")
print(paste("The number of unique values: ",length(unique(lessThan6))))
```

Add 0 to the end for all entries

#### Transform the 5 digit numbers to match the Find-A-Code format

```{r}
diag0<-c("250.7","250.6","250.4","250.8","250.1","250.2","250.3","250.5","250.9")

for(i in diag0){
  MyData_str[MyData_str==i]<-paste0(i,"0")
}

# test the codes
unique(MyData_str$diag_1[startsWith(MyData_str$diag_1, diag0)])
unique(MyData_str$diag_2[startsWith(MyData_str$diag_2, diag0)])
unique(MyData_str$diag_3[startsWith(MyData_str$diag_3, diag0)])
```

### Get the 6 digit numbers

```{r}
lessThan7<-diags[nchar(diags)==6]
unique(lessThan7)
cat("\n")
print(paste("The number of unique values: ",length(unique(lessThan7))))
```

All are valid codes 


## Export Data 

```{r}
write.csv(MyData_str,row.names=FALSE, file = "formated_diags.csv")
```


## Invalid Codes 

```{r}
print("list of the codes that doesn't match with Find-A-Code:")
nonMatchingCodes

Invalid_diag1<-MyData_str$diag_1[startsWith(MyData_str$diag_1, nonMatchingCodes)]
Invalid_diag2<-MyData_str$diag_2[startsWith(MyData_str$diag_2, nonMatchingCodes)]
Invalid_diag3<-MyData_str$diag_3[startsWith(MyData_str$diag_3, nonMatchingCodes)]

Invalid_diag1<-Invalid_diag1[nchar(Invalid_diag1)==3]
Invalid_diag2<-Invalid_diag2[nchar(Invalid_diag2)==3]
Invalid_diag3<-Invalid_diag3[nchar(Invalid_diag3)==3]



InvalidDiagsEntries<-c(Invalid_diag1,Invalid_diag2,Invalid_diag3)

cat("\n")
print("Print 20 entries of the invalid codes:")
head(InvalidDiagsEntries,20)

cat("\n")
print(paste( "There are a ",length(InvalidDiagsEntries)," entries that are not valid"))
```
## Download Transformed Data

```{r warning=FALSE, message=FALSE}
library(magrittr)
readLines("formated_diags.csv") %>% 
  paste0(collapse="\n") %>% 
  openssl::base64_encode() -> encoded
```

NOTE:

When it prompt to download GIVE THE DOCUMENT A NAME with .csv as file extension

[Download formated_diags.CSV](`r sprintf('data:text/csv;base64,%s', encoded)`)


-------
